---
title: "Argentina"
authors: "Chloe, Weilin, and Gaby"
output: pdf_document
date: "2025-04-06"
---

```{r Load packages, message=FALSE}

library(readxl)
library(openxlsx)
library(dplyr)
library(lubridate)
library(ggplot2)
library(forecast)  
library(Kendall)
library(tseries)
library(outliers)
library(tidyverse)
library(smooth)
library(zoo)
library(kableExtra)
library(writexl)
library(tsibble)
library(fable)
library(fable.prophet)

```

## Importing data
```{r Import data}

# Read Arg data from Excel file
arg_pre_uncleaned <- read.xlsx(xlsxFile = './data/Argentina_precipitation.xlsx')

# Check for NAs
colSums(is.na(arg_pre_uncleaned)) # DJF has one NA

# Interpolating a value for mean
arg_pre_uncleaned$DJF[is.na(arg_pre_uncleaned$DJF)] <- mean(arg_pre_uncleaned$DJF, 
                                                                  na.rm = TRUE)
colSums(is.na(arg_pre_uncleaned))

write.csv(arg_pre_uncleaned, './data/processed_argentina_precipitation.csv', 
          row.names = FALSE)

arg_pre <- read_csv('./data/processed_argentina_precipitation.csv')

# Inspect Brazil precipitation data
head(arg_pre)
a_nvar <- ncol(arg_pre) - 1
a_nobs <- nrow(arg_pre)
```

## Data wrangling 
```{r Data wrangling }

a_monthly_data <- arg_pre %>%
  select(YEAR, JAN, FEB, MAR, APR, MAY, JUN, JUL, AUG, SEP, OCT, NOV, DEC)

a_monthly_data_long <- a_monthly_data %>%
  pivot_longer(cols = JAN:DEC, names_to = "Month", values_to = "Precipitation")

a_monthly_data_long <- a_monthly_data_long %>%
  mutate(Month_Num = match(tolower(Month), tolower(month.abb)),
         Date = as.Date(paste(YEAR, Month_Num, "01", sep = "-")),
         .before = 1) %>%
  arrange(Date)

write.csv(a_monthly_data_long, './data/argentinamonthly_data_long.csv', row.names = FALSE)

```

## Creating a ts object
```{r Creating a ts object for monthly precipitation data}

a_ts_monthly <- ts(a_monthly_data_long$Precipitation, 
                 start = c(1901,1), frequency = 12)

# Handle outliers
a_ts_monthly_tsclean <- forecast::tsclean(a_ts_monthly_withoutliers)  #object should be a time series

autoplot(a_ts_monthly_tsclean, series="Precipitation Clean") +
  autolayer(a_ts_monthly_withoutliers, series="Precipitation Raw") +
  ylab("Monthly Precipitation in Argentina") 

# Note we only replaced missing data, algorithm did not identify any outliers for this data set.
# Even running tsclean(), no values changed.

```

## Initial Plots
```{r Initial ts plots}

autoplot(a_ts_monthly) +
  ggtitle("Monthly Precipitation Time Series") +
  xlab("Year") +
  ylab("Precipitation (mm)") +
  theme_minimal()

par(mfrow=c(1,2))
acf(a_ts_monthly, lag.max = 48, main="ACF of Monthly Precipitation")
pacf(a_ts_monthly, lag.max = 48, main="PACF of Monthly Precipitation")
# par(mfrow=c(1,1))

```
> **Interpretation:** Argentina's precipitation data seems to show seasonality judging by the equally-spaced wave pattern. There is an intense wet and dry season spike. Upon first glance, there does not seem to be a clear increasing or decreasing trend component. The ACF, which shows a wave-like pattern, indicates the possibility of what we'd certainly expect to be a seasonal component. The PACF shows a wave-like pattern, but the spikes of the pattern decay as lags progress.


## Testing and Training Data
```{r}
a_train_precipitation <- window(a_ts_monthly, end = c(2020, 12))

a_h <- 240

a_total_months <- length(ts_monthly)

a_test_precipitation <- window(a_ts_monthly, start = c(2021, 1), end = c(2024, 12))
```

## Decomposing Time Series
```{r}
# Decompose using additive
a_add_decompose_monthly <- decompose(a_ts_monthly,"additive")
plot(a_add_decompose_monthly)

# Creating de-seasonal time series
a_deseasonal_monthly <- seasadj(a_add_decompose_monthly)  

# Decompose using multiplicative
## Decomposing using multiplicative in case the magnitude of the seasonal pattern depends on the magnitude/level of the data, which is not clearly true, but let's try it. We have a question around WHETHER the magnitude of seasonality changes in relation to time.
a_mult_decompose_monthly <- decompose(a_ts_monthly, "multiplicative")
plot(a_mult_decompose_monthly)

```
```{r Detrending}

a_nvar2 <- ncol(a_monthly_data_long) - 1
a_nobs2 <- nrow(a_monthly_data_long)
t <- c(1:a_nobs2)

# Fit a linear trend to TS
a_linear_trend_model <- lm(a_monthly_data_long$Precipitation ~ t)
summary(a_linear_trend_model)

checkresiduals(a_linear_trend_model)
# The seasonal component is, of course, not being modeled. No clear trend on the residuals series.

```



## ARIMA
```{r}
a_fit_arima <- auto.arima(a_train_precipitation)
a_pre_arima <- forecast(a_fit_arima, h = 360)
checkresiduals(a_pre_arima)

plot(a_pre_arima, main = "ARIMA Forecast")
```

## STL + ETS
```{r}

a_pre_stl_ets <- stlf(a_train_precipitation, h = a_h, method = "ets")
checkresiduals(a_pre_stl_ets)
plot(a_pre_stl_ets, main = "STL + ETS Forecast")

```

## ARIMA + Fourier terms
```{r}

a_K <- 6
a_fourier_train <- fourier(a_train_precipitation, K = a_K)
a_fourier_future <- fourier(a_train_precipitation, K = a_K, h = a_h)
a_fit_fourier <- auto.arima(a_train_precipitation, xreg = a_fourier_train, seasonal = FALSE)
a_pre_fourier <- forecast(a_fit_fourier, xreg = a_fourier_future, h = a_h)

checkresiduals(a_pre_fourier)

plot(a_pre_fourier, main = "ARIMA + Fourier Forecast")
```

## TBATS
```{r}
a_fit_tbats <- tbats(a_train_precipitation)
a_pre_tbats <- forecast(a_fit_tbats, h = a_h)
plot(a_pre_tbats, main = "TBATS Forecast")

```

## Neural Network
```{r}
a_fit_nnetar <- nnetar(a_train_precipitation)
a_pre_nnetar <- forecast(a_fit_nnetar, h = a_h)
plot(a_pre_nnetar, main = "NNETAR Forecast")
```
## Comparing performance metrics across models for Argentina
```{r Comparing performance metrics across models}

a_calculate_all_metrics <- function(a_forecast_obj, a_test_data) {
  metrics <- accuracy(a_forecast_obj, a_test_data)["Test set", ]
  return(metrics)
}

a_accuracy_results <- data.frame(
  Model = c("ARIMA", "STL + ETS", "ARIMA + Fourier", "TBATS", "NNETAR"),
  ME = c(
    accuracy(a_pre_arima, a_test_precipitation)["Test set", "ME"],
    accuracy(a_pre_stl_ets, a_test_precipitation)["Test set", "ME"],
    accuracy(a_pre_fourier, a_test_precipitation)["Test set", "ME"],
    accuracy(a_pre_tbats, a_test_precipitation)["Test set", "ME"],
    accuracy(a_pre_nnetar, a_test_precipitation)["Test set", "ME"]
  ),
  RMSE = c(
    accuracy(a_pre_arima, a_test_precipitation)["Test set", "RMSE"],
    accuracy(a_pre_stl_ets, a_test_precipitation)["Test set", "RMSE"],
    accuracy(a_pre_fourier, a_test_precipitation)["Test set", "RMSE"],
    accuracy(a_pre_tbats, a_test_precipitation)["Test set", "RMSE"],
    accuracy(a_pre_nnetar, a_test_precipitation)["Test set", "RMSE"]
  ),
  MAE = c(
    accuracy(a_pre_arima, a_test_precipitation)["Test set", "MAE"],
    accuracy(a_pre_stl_ets, a_test_precipitation)["Test set", "MAE"],
    accuracy(a_pre_fourier, a_test_precipitation)["Test set", "MAE"],
    accuracy(a_pre_tbats, a_test_precipitation)["Test set", "MAE"],
    accuracy(a_pre_nnetar, a_test_precipitation)["Test set", "MAE"]
  ),
  MAPE = c(
    accuracy(a_pre_arima, a_test_precipitation)["Test set", "MAPE"],
    accuracy(a_pre_stl_ets, a_test_precipitation)["Test set", "MAPE"],
    accuracy(a_pre_fourier, a_test_precipitation)["Test set", "MAPE"],
    accuracy(a_pre_tbats, a_test_precipitation)["Test set", "MAPE"],
    accuracy(a_pre_nnetar, a_test_precipitation)["Test set", "MAPE"]
  ),
  MASE = c(
    accuracy(a_pre_arima, a_test_precipitation)["Test set", "MASE"],
    accuracy(a_pre_stl_ets, a_test_precipitation)["Test set", "MASE"],
    accuracy(a_pre_fourier, a_test_precipitation)["Test set", "MASE"],
    accuracy(a_pre_tbats, a_test_precipitation)["Test set", "MASE"],
    accuracy(a_pre_nnetar, a_test_precipitation)["Test set", "MASE"]
  )
)

a_accuracy_results_formatted <- a_accuracy_results %>%
  arrange(RMSE) %>%  # Sort by RMSE
  mutate(across(where(is.numeric), round, 3))

print(a_accuracy_results_formatted)
```

## Comparing plots of models all together
```{r Comparing plots of models all together}

autoplot(a_test_precipitation, main = "Comparing forecasts for Argentina") +
   autolayer(a_pre_arima, PI=FALSE, series="ARIMA") +
   autolayer(a_pre_stl_ets, PI=FALSE, series="STL+ETS") +
   autolayer(a_pre_fourier, PI=FALSE, series="ARIMA + Fourier") +
   autolayer(a_pre_tbats,PI=FALSE, series="TBATS") +
   autolayer(a_pre_nnetar,PI=FALSE, series="NN") +
  xlab("Year") + 
  ylab("Precipitation (mm)") + 
  guides(colour = guide_legend(title = "Forecast")) 

```

```{r Comparing first 60 yrs and second 60 yrs}

## Our data begins at 1901--we are lucky to be working with records from over more than 100 years.
## Let's break the data into two parts, 1901-1962 and 1963-2024.
## We want to then compare the mean of the high points and the mean of the low points to see if the mangitudes of the extremes are changing.

```


