---
title: "Brazil"
authors: "Chloe, Weilin, and Gabby"
output: html_document
date: "2025-04-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=60), tidy=TRUE) 
```

```{r package, message=FALSE}
library(readxl)
library(openxlsx)
library(dplyr)
library(lubridate)
library(ggplot2)
library(forecast)  
library(Kendall)
library(tseries)
library(outliers)
library(tidyverse)
library(smooth)
library(zoo)
library(kableExtra)
library(writexl)
library(tsibble)
library(fable)
library(fable.prophet)

```

## Importing Data
```{r}

#importing data
Colombia_pre_uncleaned <- read_excel("../data/Colombia_precipitation.xlsx")   # I changed the to the relative address 

colSums(is.na(Colombia_pre_uncleaned))
Colombia_pre_uncleaned$DJF[is.na(Colombia_pre_uncleaned$DJF)] <- mean(Colombia_pre_uncleaned$DJF, na.rm = TRUE)
colSums(is.na(Colombia_pre_uncleaned))

write.csv(Colombia_pre_uncleaned, "../data/processed_colombia_precipitation.csv", row.names = FALSE)

colombia_pre <- read_csv("../data/processed_colombia_precipitation.csv")

head(colombia_pre)
nvar <- ncol(colombia_pre) - 1
nobs <- nrow(colombia_pre)

```


## Initial Plots & Time Series - Monthly
```{r}

monthly_data <- colombia_pre %>%
  select(YEAR, JAN, FEB, MAR, APR, MAY, JUN, JUL, AUG, SEP, OCT, NOV, DEC)

monthly_data_long <- monthly_data %>%
  pivot_longer(cols = JAN:DEC, names_to = "Month", values_to = "Precipitation")

monthly_data_long <- monthly_data_long %>%
  mutate(Month_Num = match(tolower(Month), tolower(month.abb)),
         Date = as.Date(paste(YEAR, Month_Num, "01", sep = "-")),
         .before = 1) %>%
  arrange(Date)

write.csv(monthly_data_long, "../data/colombiamonthly_data_long.csv", row.names = FALSE)

ts_monthly <- ts(monthly_data_long$Precipitation, 
                 start = c(1901,1), frequency = 12)

```

## Initial Plots
```{r}
autoplot(ts_monthly) +
  ggtitle("Monthly Precipitation Time Series") +
  xlab("Year") +
  ylab("Precipitation (mm)") +
  theme_minimal()

par(mfrow=c(1,2))
acf(ts_monthly, lag.max = 48, main="ACF of Monthly Precipitation")
pacf(ts_monthly, lag.max = 48, main="PACF of Monthly Precipitation")
par(mfrow=c(1,1))
```
## Testing and Training Data
```{r}

train_precipitation <- window(ts_monthly, end = c(2020, 12))

h <- 240

total_months <- length(ts_monthly)

test_precipitation <- window(ts_monthly, start = c(2021, 1), end = c(2024, 12))

```

## Decomposing Time Series
```{r}

# Decompose
decompose_monthly <- decompose(ts_monthly,"additive")
plot(decompose_monthly)

#Creating non-seasonal time series
deseasonal_monthly <- seasadj(decompose_monthly)  

```
## ARIMA
```{r}

fit_arima <- auto.arima(train_precipitation)
pre_arima <- forecast(fit_arima, h = h)

plot(pre_arima, main = "ARIMA Forecast")

```

## STL + ETS
```{r}

pre_stl_ets <- stlf(train_precipitation, h = h, method = "ets")
plot(pre_stl_ets, main = "STL + ETS Forecast")

```

## ARIMA + Fourier terms
```{r}

K <- 6
fourier_train <- fourier(train_precipitation, K = K)
fourier_future <- fourier(train_precipitation, K = K, h = h)
fit_fourier <- auto.arima(train_precipitation, xreg = fourier_train, seasonal = FALSE)
pre_fourier <- forecast(fit_fourier, xreg = fourier_future, h = h)

plot(pre_fourier, main = "ARIMA + Fourier Forecast")

```

## TBATS
```{r}

fit_tbats <- tbats(train_precipitation)
pre_tbats <- forecast(fit_tbats, h = h)
plot(pre_tbats, main = "TBATS Forecast")

```

## Neural Network
```{r}

fit_nnetar <- nnetar(train_precipitation)
pre_nnetar <- forecast(fit_nnetar, h = h)
plot(pre_nnetar, main = "NNETAR Forecast")

```


```{r}

print(data.frame(
  Model = c("ARIMA", "STL + ETS", "ARIMA + Fourier", "TBATS", "NNETAR"),
  MAE = c(
    accuracy(fc_arima, test_precipitation)["Test set", "MAE"],
    accuracy(fc_stl_ets, test_precipitation)["Test set", "MAE"],
    accuracy(fc_fourier, test_precipitation)["Test set", "MAE"],
    accuracy(fc_tbats, test_precipitation)["Test set", "MAE"],
    accuracy(fc_nnetar, test_precipitation)["Test set", "MAE"]
  ),
  RMSE = c(
    accuracy(fc_arima, test_precipitation)["Test set", "RMSE"],
    accuracy(fc_stl_ets, test_precipitation)["Test set", "RMSE"],
    accuracy(fc_fourier, test_precipitation)["Test set", "RMSE"],
    accuracy(fc_tbats, test_precipitation)["Test set", "RMSE"],
    accuracy(fc_nnetar, test_precipitation)["Test set", "RMSE"]
  ),
  MAPE = c(
    accuracy(fc_arima, test_precipitation)["Test set", "MAPE"],
    accuracy(fc_stl_ets, test_precipitation)["Test set", "MAPE"],
    accuracy(fc_fourier, test_precipitation)["Test set", "MAPE"],
    accuracy(fc_tbats, test_precipitation)["Test set", "MAPE"],
    accuracy(fc_nnetar, test_precipitation)["Test set", "MAPE"]
  )
))

```




OLD

```{r}
msts_monthly <- msts(monthly_data_long$Precipitation, 
                     seasonal.periods =c(12, 365.25),
                 start = c(min(monthly_data_long$YEAR), 1))

autoplot(msts_monthly) +
  ggtitle("Monthly Precipitation Time Series") +
  xlab("Year") +
  ylab("Precipitation (mm)") +
  theme_minimal()

par(mfrow=c(1,2))
acf(msts_monthly, lag.max = 48, main="ACF of Monthly Precipitation")
pacf(msts_monthly, lag.max = 48, main="PACF of Monthly Precipitation")
par(mfrow=c(1,1))

```

## Testing and Training Data - Monthly
```{r}

# Training data: 1901-2020
training_data_monthly <- window(msts_monthly, end=c(2020))

# Test data: 2021-2024
test_data_monthly <- window(ts_monthly, start=c(2021))

autoplot(training_data_monthly)
autoplot(test_data_monthly)

```

## Decomposing Time Series
```{r}

# Decompose
decompose_monthly <- decompose(ts_monthly,"additive")
plot(decompose_monthly)

#Creating non-seasonal time series
deseasonal_monthly <- seasadj(decompose_monthly)  

```

## ARIMA forecast
```{r}

SARIMA_autofit <- auto.arima(ts_monthly)
print(SARIMA_autofit)

SARIMA_forecast <- forecast(object = SARIMA_autofit, h = 240)
plot(SARIMA_forecast)

```

## STL + ETS
```{r}
#Fit and forecast STL + ETS model to data
ETS_fit <-  stlf(training_data_monthly,h=365)

#Plot foresting results
autoplot(ETS_fit) + ylab("Precipitation (mm)")

#Plot model + observed data
autoplot(msts_monthly) +
  autolayer(ETS_fit, series="STL + ETS",PI=FALSE) +
  ylab("Precipitation (mm)")

```

## ARIMA Fourier
```{r}
#Fit arima model with fourier terms as exogenous regressors
# seasonal = FALSE is the same as P=D=Q=0
# play with K by changing it to K=c(2,2), K=c(2,4), K=c(2,6), etc. The higher teh K the longer it will take to converge, because R will try more models.


K <- c(2, 3) 

ARIMA_Four_fit <- auto.arima(training_data_monthly, xreg = fourier(training_data_monthly, K = K))

ARIMA_Four_for <- forecast(ARIMA_Four_fit,
                           xreg = fourier(training_data_monthly, K = K, h = 12),
                           h = 12)

autoplot(ARIMA_Four_for) + ylab("Precipitation (mm)")

#Plot model + observed data
autoplot(msts_monthly) +
  autolayer(ARIMA_Four_for, series="ARIMA_FOURIER",PI=FALSE) +
  ylab("Precipitation (mm)")

```
training_data_monthly_ts <- ts(training_data_monthly, frequency = 12)

ARIMA_Four_fit <- auto.arima(training_data_monthly, 
                             seasonal=FALSE, 
                             lambda=0,
                             xreg=fourier(training_data_monthly_ts, 
                                          K=2)
                             )

#Forecast with ARIMA fit
#also need to specify h for fourier terms
ARIMA_Four_for <- forecast(ARIMA_Four_fit,
                           xreg=fourier(training_data_monthly_ts,
                                        K=2,
                                        h = 12),
                           h = 12
                           ) 


autoplot(ARIMA_Four_for) + ylab("Precipitation (mm)")

#Plot model + observed data
autoplot(msts_monthly) +
  autolayer(ARIMA_Four_for, series="ARIMA_FOURIER",PI=FALSE) +
  ylab("Precipitation (mm)")



## TBATS
```{r}
# TBATS can take time to fit
TBATS_fit <- tbats(training_data_monthly)

TBATS_for <- forecast(TBATS_fit, h=365)

#Plot foresting results
autoplot(TBATS_for) +
  ylab("Precipitation (mm)") 

#Plot model + observed data
autoplot(training_data_monthly) +
  autolayer(TBATS_for, series="TBATS",PI=FALSE)+
  ylab("Precipitation (mm)") 
```

```{r}

#You can play with the different values for p and P, you can also use xreg with Fourier term to model the multiple seasonality

#NN_fit <- nnetar(ts_act_power_daily_train,p=1,P=1)
NN_fit <- nnetar(training_data_monthly,
                 p=1,
                 P=0,
                 xreg=fourier(training_data_monthly, K=2))

#NN_for <- forecast(NN_fit, h=365) 
NN_for <- forecast(NN_fit, 
                   h=365,
                   xreg=fourier(training_data_monthly, 
                                          K=2,h=365))


#Plot foresting results
autoplot(NN_for) +
  ylab("Precipitation (mm)") 


#Plot model + observed data
autoplot(ts_monthly) +
  autolayer(NN_for, series="Neural Network",PI=FALSE)+
  ylab("Precipitation (mm)") 

```
