---
title: "Old Code Colombia"
output: html_document
date: "2025-04-06"
---

OLD

```{r}
msts_monthly <- msts(monthly_data_long$Precipitation, 
                     seasonal.periods =c(12, 365.25),
                 start = c(min(monthly_data_long$YEAR), 1))

autoplot(msts_monthly) +
  ggtitle("Monthly Precipitation Time Series") +
  xlab("Year") +
  ylab("Precipitation (mm)") +
  theme_minimal()

par(mfrow=c(1,2))
acf(msts_monthly, lag.max = 48, main="ACF of Monthly Precipitation")
pacf(msts_monthly, lag.max = 48, main="PACF of Monthly Precipitation")
par(mfrow=c(1,1))

```

## Testing and Training Data - Monthly
```{r}

# Training data: 1901-2020
training_data_monthly <- window(msts_monthly, end=c(2020))

# Test data: 2021-2024
test_data_monthly <- window(ts_monthly, start=c(2021))

autoplot(training_data_monthly)
autoplot(test_data_monthly)

```

## Decomposing Time Series
```{r}

# Decompose
decompose_monthly <- decompose(ts_monthly,"additive")
plot(decompose_monthly)

#Creating non-seasonal time series
deseasonal_monthly <- seasadj(decompose_monthly)  

```

## ARIMA forecast
```{r}

SARIMA_autofit <- auto.arima(ts_monthly)
print(SARIMA_autofit)

SARIMA_forecast <- forecast(object = SARIMA_autofit, h = 240)
plot(SARIMA_forecast)

```

## STL + ETS
```{r}
#Fit and forecast STL + ETS model to data
ETS_fit <-  stlf(training_data_monthly,h=365)

#Plot foresting results
autoplot(ETS_fit) + ylab("Precipitation (mm)")

#Plot model + observed data
autoplot(msts_monthly) +
  autolayer(ETS_fit, series="STL + ETS",PI=FALSE) +
  ylab("Precipitation (mm)")

```

## ARIMA Fourier
```{r}
#Fit arima model with fourier terms as exogenous regressors
# seasonal = FALSE is the same as P=D=Q=0
# play with K by changing it to K=c(2,2), K=c(2,4), K=c(2,6), etc. The higher teh K the longer it will take to converge, because R will try more models.


K <- c(2, 3) 

ARIMA_Four_fit <- auto.arima(training_data_monthly, xreg = fourier(training_data_monthly, K = K))

ARIMA_Four_for <- forecast(ARIMA_Four_fit,
                           xreg = fourier(training_data_monthly, K = K, h = 12),
                           h = 12)

autoplot(ARIMA_Four_for) + ylab("Precipitation (mm)")

#Plot model + observed data
autoplot(msts_monthly) +
  autolayer(ARIMA_Four_for, series="ARIMA_FOURIER",PI=FALSE) +
  ylab("Precipitation (mm)")

```
training_data_monthly_ts <- ts(training_data_monthly, frequency = 12)

ARIMA_Four_fit <- auto.arima(training_data_monthly, 
                             seasonal=FALSE, 
                             lambda=0,
                             xreg=fourier(training_data_monthly_ts, 
                                          K=2)
                             )

#Forecast with ARIMA fit
#also need to specify h for fourier terms
ARIMA_Four_for <- forecast(ARIMA_Four_fit,
                           xreg=fourier(training_data_monthly_ts,
                                        K=2,
                                        h = 12),
                           h = 12
                           ) 


autoplot(ARIMA_Four_for) + ylab("Precipitation (mm)")

#Plot model + observed data
autoplot(msts_monthly) +
  autolayer(ARIMA_Four_for, series="ARIMA_FOURIER",PI=FALSE) +
  ylab("Precipitation (mm)")



## TBATS
```{r}
# TBATS can take time to fit
TBATS_fit <- tbats(training_data_monthly)

TBATS_for <- forecast(TBATS_fit, h=365)

#Plot foresting results
autoplot(TBATS_for) +
  ylab("Precipitation (mm)") 

#Plot model + observed data
autoplot(training_data_monthly) +
  autolayer(TBATS_for, series="TBATS",PI=FALSE)+
  ylab("Precipitation (mm)") 
```

```{r}

#You can play with the different values for p and P, you can also use xreg with Fourier term to model the multiple seasonality

#NN_fit <- nnetar(ts_act_power_daily_train,p=1,P=1)
NN_fit <- nnetar(training_data_monthly,
                 p=1,
                 P=0,
                 xreg=fourier(training_data_monthly, K=2))

#NN_for <- forecast(NN_fit, h=365) 
NN_for <- forecast(NN_fit, 
                   h=365,
                   xreg=fourier(training_data_monthly, 
                                          K=2,h=365))


#Plot foresting results
autoplot(NN_for) +
  ylab("Precipitation (mm)") 


#Plot model + observed data
autoplot(ts_monthly) +
  autolayer(NN_for, series="Neural Network",PI=FALSE)+
  ylab("Precipitation (mm)") 

```